---
title: "covid 19"
author: "Michael Abdalla"
date: "7/15/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata$sex <- as.factor(kRdata$sex)
kRdata$result <- as.factor(kRdata$result)
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
barchartGC(~result + sex,data=kRdata)
barchartGC(~result + sex,data=kRdata, type="percent")

```

the bar shows that the male deaths related to covid make up to 60% of the total deaths and the females make the other 40%

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

the numerical results show us that out of 100, males make up to 64% of covid related death compared to the females at a percentage of 36% as well as this data shows that males are twice likely to get affected with covid than the women

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

fisher's exact test for count data shows that males are 2.3 times more likely to die out of covid compared to the females as will as the p-value that is shown here is very tiny number which will make us reject the null hypothesis and the states that have covid deaths are independent of sex.

## Conclusions

